Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-10-01 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:10:01	21:39:48	(Thu 1.Oct)

\begin{figure*}[t]\centering%
\includegraphics[width=\linewidth]{gfx/mvdr_selected_windows_holmengraa.pdf}%
\caption{\emph{Determining LCA window type}: MVDR image with typical frequency responses for windows used in various pixel regions.\newline
\emph{Left:}\hfill
\parbox[t]{.95\linewidth}{MVDR sectorscan image of the oiltanker Holmengraa, using data collected from the 32 element HISAS1030 sonar attached to the HUGIN AUV. MVDR was run with $K=1$ temporal oversampling, $L=16$ subarray size, $d=1\%$ diagonal loading and 8 times lateral oversampling.}\protect\\\hspace{\textwidth}
\emph{Right:}\hfill
\parbox[t]{.95\linewidth}{MVDR window responses computed from 40x40 pixel groups in the shadow, highlight and speckle region of the Holmengraa image. The responses are overlayed each other. The dashed red lines at -13dB and -43dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3dB in highlight. The phase varies mostly in highlight to allow the steering there.} }\label{mvdr_selected_windows}
\end{figure*}

@@

Observe from \Fig{mvdr_selected_windows} that MVDR seems to prefer symmetric window responses, even if it is free to choose non-symmetric ones. The symmetry is most predominant in shadow and speckle regions, while in highlight the windows are approximately symmetric within the illuminated sector of the seafloor. We have observed this symmetry to be a side-effect of the averaging steps needed to build the sample covariance matrix, which is required to allow MVDR to run in near snapshot mode in an active system. The peak sidelobe levels of the MVDR windows are mostly between -13\,dB and -43\,dB. This corresponds to that of the rectangular and Hamming window, or a Kaiser window with $\beta=0$ and $\beta=2$, respectively.

@@

The images in \Fig{oversampling_mosaic_bounds} much improved sidelobe suppression with a an upper for bound for $\beta$ of 2, but it improved further as this bound was adjusted up to 5. No noticeable difference were observed for higher values than 5. The images also demonstrate that increasing the steering $\phi$ improves the resolution of the strong scatterers. For this sonar the Rayleigh resolution is $\Delta=0.0125$ radians (0.72$^\circ$), thus a point scatterer imaged at 19\,m with a rectangular DAS would have a lateral FWHM of $0.125\cdot19\,\text{m}\cdot0.89=21\,\text{cm}$. This is close to the FWHM of the anchor in the DAS image in \Fig{oversampling_mosaic_bounds}, which means that the acoustic fingerprint of the anchor is similar to a point source. This suggests the anchor reflections to be specular and that we view its rounded side. As we increase steering to either 50\% or 100\% of the -3dB width, the FWHM drops to 73\% or 53\% to that of the rectangular window, respectively.

@@

LCA iterates through a set of windows in search of the one that minimizes beamformer output power. It can be viewed as either a reduced window solution space version of  MVDR, or as an extension to DAS. It creates images similar to MVDR, but at a fraction of the computational cost. It is also inherently robust and insensitive to the particular set of windows it selects from, supported by the observation that LCA can be used with either Kaiser or trigonometric windows and produce near identical images.

>>>>

\begin{figure*}[t]\centering%
\includegraphics[width=\linewidth]{gfx/mvdr_selected_windows_holmengraa.pdf}%
\caption{\emph{Determining LCA window type}: MVDR image with typical frequency responses for windows used in various pixel regions.\newline
\emph{Left:}\hfill
\parbox[t]{.95\linewidth}{MVDR sectorscan image of the oiltanker Holmengraa, using data collected by the 32 element HISAS1030 sonar attached to the HUGIN AUV.}\protect\\\hspace{\textwidth}
\emph{Right:}\hfill
\parbox[t]{.95\linewidth}{MVDR window responses computed from 40x40 pixel groups in the shadow, highlight and speckle region of the image. The responses are overlayed each other and the amount of overlap is colored with a logarithmic scale. The dashed red lines at -13dB and -43dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3dB in highlight. The phase varies mostly in highlight to allow the steering there.} }\label{mvdr_selected_windows}
\end{figure*}

@@

Observe from \Fig{mvdr_selected_windows} that MVDR seems to prefer symmetric window responses, even if it is free to choose non-symmetric ones. The symmetry is most predominant in shadow and speckle regions, while in highlight the windows are approximately symmetric within the illuminated sector of the seafloor. We have observed this symmetry to be a side-effect of the averaging steps needed to build the sample covariance matrix, which is required to allow MVDR to run in near snapshot mode in an active system. The peak sidelobe levels of the MVDR windows are mostly between -13\,dB and -43\,dB. This corresponds to that of an unsteered rectangular and Hamming window, or an unsteered Kaiser window with $\beta=0$ and $\beta=2$, respectively.

@@

The LCA images in \Fig{oversampling_mosaic_bounds} demonstrate the effect of adjusting the upper bound of $\beta$ and $\phi$. As suggested by \Fig{mvdr_selected_windows} we attempted first to use $\beta\in[0,2]$. While this resulted in much improved sidelobe suppression, there was also a slight improvement up to $\beta\in[0,5]$. This is likely needed to compensate for the increased sidelobe levels caused by steering. No noticeable difference were observed for higher values than 5. The images also demonstrate that increasing the steering $\phi$ improves the resolution of the strong scatterers. For this sonar the Rayleigh resolution is $\Delta=0.0125$ radians (0.72$^\circ$), thus a point scatterer imaged at 19\,m with a rectangular DAS would have a lateral FWHM of $0.125\cdot19\,\text{m}\cdot0.89=21\,\text{cm}$. This is close to the FWHM of the anchor in the DAS image in \Fig{oversampling_mosaic_bounds}, which means that the acoustic fingerprint of the anchor is similar to a point source. This suggests the anchor reflections to be specular and that we view its rounded side. As we increase steering to either 50\% or 100\% of the -3dB width, the FWHM drops to 73\% or 53\% to that of the rectangular window, respectively.

@@

LCA iterates through a set of windows in search of the one that minimizes the noise and interference power in each image pixel. It can be viewed as either a reduced solution space version of MVDR, or as an multi-taper extension to DAS. It creates images similar to MVDR, but at a fraction of the computational cost. It is also inherently robust and insensitive to the particular set of windows it selects from, supported by the observation that LCA can be used with either Kaiser or trigonometric windows and produce near identical images.

We have observed that when MVDR is constrained to work in an active system it tends to create near symmetric responses with maximum sidelobe levels between that of the rectangular and Hamming window, or Kaiser windows with $\beta=0$ and $\beta=2$, respectively.
